BFS search in miniKanren

### 2019-02-28 Paper/paper.tex
update

The \emph{cost} of an answer is the number of relation applications needed to find the answer. We use the miniKanren relation \texttt{repeato} in Fig.~\ref{def-repeato} to demonstrate the cost of answers. We borrow this idea from Silvija Seres's work [*]. \texttt{repeato} relates \texttt{x} with a list whose elements are all \texttt{x}s. The run expression uses \texttt{repeato} to generate 4 lists whose elements are all \texttt{*}s. The order of the answers reflects the order miniKanren discovers them. The leftmost answer is discovered first. The order here is not suprising: to generate the answer \texttt{'()}, miniKanren needs to apply \texttt{repeato} only once. And it needs more applications of \texttt{repeato} to find the later answers that are more complicated. In this example, the cost of each answer is the same as one more than the number of \texttt{*}s. No answer of zero cost exists.

@@

Now we change the search strategy and optimize the system. The whole process is completed in three steps, corresponding to 4 versions of miniKanren. The initial version, \texttt{mk-0} is exactly the miniKanren in \emph{The Reasoned Schemer, 2nd Edition}.

@@

In mk-0, search spaces are represented by streams of answers. Streams can be finite or infinite. Finite streams are just lists. And infinite streams are improper lists, whose last \texttt{cdr} is a thunk returning another stream. We call the \texttt{car}s the \emph{mature} part, and the last \texttt{cdr} the \emph{immature} part.

Streams are cost respective when they are initially constructed by \texttt{==}. However, \texttt{append-inf}, a composer of stream, breaks cost respectiveness. By changing its definition, we can make the whole search strategy cost respective.

(explain mk-0's append-inf)

@@

@@

@@

>>>>

The \emph{cost} of an answer is the number of relation applications needed to find the answer. We use the miniKanren relation \texttt{repeato} in Fig.~\ref{def-repeato} to demonstrate the concept. We borrow this idea from Silvija Seres's work [*]. \texttt{repeato} relates a term \texttt{x} with a list whose elements are all \texttt{x}s. The \texttt{run} expression uses \texttt{repeato} to generate 4 lists whose elements are all \texttt{*}s. The order of the answers reflects the order miniKanren discovers them. The leftmost answer is discovered first. The order here is not suprising: to generate the answer \texttt{'()}, miniKanren needs to apply \texttt{repeato} only once. And it needs more applications of \texttt{repeato} to find the later answers that are more complicated. In this example, the cost of each answer is the same as one more than the number of \texttt{*}s. No answer of zero cost exists.

@@

Now we change the search strategy and optimize the system. The whole process is completed in three steps, corresponding to 4 versions of miniKanren. The initial version, \texttt{mk-0}, is exactly the miniKanren in \emph{The Reasoned Schemer, 2nd Edition}.

@@

In \texttt{mk-0} and \texttt{mk-1}, search spaces are represented by streams of answers. Streams can be finite or infinite. Finite streams are just lists. And infinite streams are improper lists, whose last \texttt{cdr} is a thunk returning another stream. We call the \texttt{car}s the \emph{mature} part, and the last \texttt{cdr} the \emph{immature} part.

Streams are cost respective when they are initially constructed by \texttt{==}. However, the \texttt{mk-0} version of \texttt{append-inf} (Fig.~\ref{append-inf-0}), a composer of stream, can break cost respectiveness. when its first argument, \texttt{s-inf} is infinite. The resulting mature part contains only the mature part of \texttt{s-inf}. The whole \texttt{t-inf} goes to the resulting immature part -- the newly constructed thunk.

@@

The \texttt{mk-1} version of \texttt{append-inf} (Fig.~\ref{append-inf-1}) restores cost-respectiveness by combining the mature parts in the fashion of \texttt{append}. \texttt{append-inf} calls its helper immediately, with the first argument, \texttt{s?}, set to \texttt{\#{}t}, which means \texttt{s-inf} in the helper is the \texttt{s-inf} in the driver. Two streams are swapped in the third \texttt{cond} clause, where \texttt{s?} is also changed accordingly.

@@

\texttt{mk-1} is not efficient in two aspects. \texttt{append-inf} need to copy all \texttt{cons} cells of two input stream when the first one is infinite. Besides, \texttt{mk-1} generates answers of the same cost at once, even when only a small portion is queried. We solves the two problems in the next two subsections.

@@

(eqv to Seres's BFS ?)
